---
title: "Assignment2"
author: "Prithvi Abhishetty"
date: "2022-11-03"
output: html_document
---

## 1.1 (Q1)

```{r, include=FALSE}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

```{r}
head(Hawks)
```

```{r}
hSF <- Hawks %>% 
  filter(Species=='RT' & Weight>=1000) %>% 
  select(Wing, Weight, Tail)
head(hSF, 5)
```

## 1.1 (Q2)

3 variables (Wing, weight, tail) = features

examples/observations/cases = number of rows
```{r}
hSF %>% nrow()
```

## 1.2 (Q1)

```{r}
hSF %>% arrange(Wing) %>% head(5)
```

## 1.3 (Q1)

```{r}
species_code <- c('CH','RT','SS')
species_name_full <- c('Cooper\'s','Red-tailed','Sharp-shinned')
hawkSpeciesNameCodes <- data.frame(species_code,species_name_full)
print(hawkSpeciesNameCodes)
```

## 1.3 (Q2)

```{r}
hawksFullName <- Hawks %>% 
  left_join(hawkSpeciesNameCodes %>% 
              rename(Species=species_code)) %>% 
  select(-Species) %>%
  rename(Species=species_name_full)
```

## 1.3 (Q3)

```{r}
hawksFullName %>% select(Species,Wing,Weight) %>% head(7)
```
Makes no difference what joining method you use because both data frames use the same set of species codes.

Only matters if they have different entries (rows)

## 1.4 (Q1)

```{r}
hawksWithBMI <- Hawks %>% 
  mutate(bird_BMI=1000*Weight/Wing^{2}) %>% 
  select(Species,bird_BMI) %>%
  arrange(-bird_BMI)
hawksWithBMI %>% head(8)
```

## 1.5 (Q1)

```{r}
hawksFullName %>% 
  group_by(Species) %>% 
  summarise(num_rows=n(),mn_wing=mean(Wing, na.rm=1),md_wing=median(Wing, na.rm=1),t_mn_wing=mean(Wing, na.rm=1, trim=0.1),b_wt_ratio=max(Wing/Tail, na.rm=1))
```
  
## 1.5 (Q2)

```{r}
hawksFullName %>%
  select(Species, Wing, Weight, Culmen, Hallux, Tail, StandardTail, Tarsus, Crop) %>%
  group_by(Species) %>% 
  summarise(across(everything(), ~sum(is.na(.x))))
```

## 2.1 (Q1)

You would need a Tidyverse library.

## 2.1 (Q2)

```{r}
impute_by_median<-function(x){
  md<-median(x,na.rm=1) # first compute the median of x
  impute_f<-function(z){ # coordinate-wise imputation
  if(is.na(z)){
    return(md) # if z is na replace with median
    }else{
      return(z) # otherwise leave in place
    }
  }
return(map_dbl(x,impute_f)) # apply the map function to impute across vector
}
v<-c(1,2,NA,4)
impute_by_median(v)
```
## 2.1 (Q3)

```{r}
x <- seq(from=0, to=10, by=0.1)
y <- c(5*x+1)
df_xy <- data.frame(x,y)
df_xy %>% head(5)
```
```{r}
df_xy %>% mutate(z=map2_dbl(x,y,~.x+.y))
df_xy %>% head(5)
```

## 2.1 (Q4)

```{r}
sometimes_missing <- function(index,value){
  if (index %% 5 == 0){
    return(NA)
  }else{
    return(value)
  } 
}
sometimes_missing(14,25)
sometimes_missing(15,25)
```

```{r}
df_xy_missing <- df_xy %>%
  mutate(y=map2_dbl(row_number(),y,sometimes_missing))

df_xy_missing %>% head(10)
```

## 2.1 (Q5)

```{r}
df_xy_imputed <- df_xy_missing %>%
  mutate(y=impute_by_median(y))

df_xy_imputed %>% head()
```  

## 2.2 (Q1)

```{r}
library(readxl)
folder_path <- "C:/Users/prith/OneDrive/Documents/DataScienceMSc/SCEM/SCEM_R_Projects/"
file_name <- "HockeyLeague.xlsx"
file_path <- paste(folder_path,file_name,sep="")
wins_data_frame <- read_excel(file_path,sheet="Wins")

wins_data_frame %>%
  select(1:5) %>%
  head(3)
```
Not usually considered as tidy data because the number of games won, total number of games and the year are all contained in the same column therefore each column corresponds to multiple variables.

```{r}
w_l_narrow <- function(w_or_l){
  return(
    read_excel(file_path,sheet=w_or_l) %>%
      rename(Team=...1) %>% # rename first column title from ...1 to Team
      pivot_longer(!Team, names_to="Year",values_to="val") %>% # reshape all columns except Team, such that the titles become a column called Year and the values become a column called val.
      mutate(Year=as.integer(Year)) %>% # change Year column from character to integer
      separate(col=val,into=c(w_or_l,"Total"),sep=" of ",convert = TRUE) # separate val column to Wins and Total and the values of the column are separated by splitting on the " of " part of the values.
  )
}
wins_tidy <- w_l_narrow(w_or_l="Wins") # Take Wins sheet of excel document
wins_tidy %>% dim() # gives dimensions of data frame
wins_tidy %>% head(5)

```

## 2.2 (Q2)

```{r}
losses_tidy <- w_l_narrow(w_or_l = "Losses")
losses_tidy %>% head(5)
```

## 2.2 (Q3)

```{r}
hockey_df <- wins_tidy %>% 
  inner_join(losses_tidy) %>% 
  mutate(Draws=Total-Wins-Losses, across(starts_with(c("Wins","Losses","Draws")),~.x/Total, .names="{.col}_rt"))
hockey_df %>% dim()
hockey_df %>% head(5)
```

## 2.2 (Q4)

```{r}
hockey_df %>%
select(-Wins,-Draws,-Losses) %>%
group_by(Team) %>%
summarise(across(c("Wins_rt","Losses_rt","Draws_rt"),
list(md=median,mn=mean),.names="{substring(.col,1,1)}_{.fn}")) %>%
arrange(desc(W_md))
  
```

## 3. (Q1)

```{r}
ggplot(data=filter(wins_tidy, Team=="Ducks"), aes(x=Wins)) + xlab("wins") + ylab("Count") + geom_histogram(binwidth = 3)

```

## 3. (Q2)

```{r}
ggplot(data=filter(wins_tidy, Team=="Ducks"), aes(x=Wins)) + xlab("wins") + ylab("Count") + geom_density(adjust=0.5)
ggplot(data=filter(wins_tidy, Team=="Ducks"), aes(x=Wins)) + xlab("wins") + ylab("Count") + geom_density(adjust=2)
```

The second plot curve is a smoother version of the first plot because the adjust argument in geom_density() adjusts the bandwidth of the density plot. Larger bandwidth = smoother function.

## 3. (Q3)

```{r}
wins_teams <- wins_tidy %>% pivot_wider(names_from = Team, values_from = Wins) %>% select(-Total)
wins_teams
ggplot(data = wins_teams, aes(y=Eagles, x=Ducks)) + xlab("Ducks") + ylab("Eagles") + geom_point()
```
