---
title: "Assignment 6"
author: "Prithvi Abhishetty"
date: "2022-11-21"
output: html_document # you can change to other output format if you want
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 1. Continuous random variables and limit laws

## 1.1 (Q1)

## 1.1 (Q2)
```{r}
set.seed(0)
n <- 1000
sample_X <- data.frame(U=runif(n)) %>%
mutate(X=case_when(
(0<=U)&(U<0.25)~3,
(0.25<=U)&(U<0.5)~10,
(0.5<=U)&(U<=1)~0)) %>%
pull(X)
```

## 1.1 (Q3)

```{r}
sample_X_0310 <- function(alpha,beta,n){
  sample_X <- data.frame(U=runif(n)) %>%
    mutate(X=case_when(
      (0<=U)&(U<alpha)~3,
      (alpha<=U)&(U<alpha+beta)~10,
      (alpha+beta<=U)&(U<=1)~0
    )) %>%
    pull(X)
  return(sample_X)
}
```

## 1.1 (Q4)

```{r}
sample_X_0310(1/2,1/10,1000) %>% mean()

eX <- (1/2)*(3)+(1/10)*(10)
print(eX)
```
The sample mean is close to the population mean because the number of trials is large, hence follows the law of large numbers.

## 1.1 (Q5)

```{r}
sample_X_0310(1/2,1/10,1000) %>% var()

eX2 <- (1/2)*(3^2)+(1/10)*(10^2)

variance <- eX2 - (eX)^2
print(variance)
```
The sample variance is close to the population variance, and the number of trials are large, hence this follows the law of large numbers.

## 1.1 (Q6)

```{r}
n = 100
alpha = 1/10
samples <- data.frame(beta = seq(0,9/10,0.01)) %>% 
  mutate(sample_X = map(beta, ~sample_X_0310(alpha,.x,n))) %>%
  mutate(samplemean = map_dbl(sample_X, mean)) %>%
           mutate(Expectation = 3*alpha + 10*beta)
           
```
## 1.1 (Q7)

```{r}
df <- samples %>% pivot_longer(cols=c(samplemean, Expectation), names_to = 'name', values_to = 'value')

ggplot(df, aes(x=beta, y=value, color=name)) + geom_point() + theme_bw()
```


## 1.2 (Q2)

```{r}
my_cdf_exp<-function(x,lambda){
  if (x<0){return (0)}
  return(1-exp(-lambda*x))
}
lambda<-1/2
map_dbl(.x=seq(-1,4), .f=~my_cdf_exp(x=.x,lambda=lambda))

test_inputs <- seq(-1,10,0.1)
my_cdf_output <- map_dbl(.x=test_inputs, .f=~my_cdf_exp(x=.x,lambda=lambda))
inbuilt_cdf_output <- map_dbl(.x=test_inputs,.f=~pexp(q=.x,rate=lambda))
all.equal(my_cdf_output,inbuilt_cdf_output)
```
## 1.2 (Q3)

```{r}
my_quantile_exp<-function(p, lambda){
  if (p<=0){return(-Inf)}
  return(log(1-p)/-lambda)
}

test_inputs <- seq(0.01, 0.99, 0.01)
my_quantile_output <- map_dbl(.x=test_inputs, .f=~my_quantile_exp(p=.x,lambda=lambda))
inbuilt_quantile_output <- map_dbl(.x=test_inputs,.f=~qexp(p=.x,rate=lambda))
all.equal(my_quantile_output,inbuilt_quantile_output)

```

## 1.3 (Q2)

```{r}
binom_df <- data.frame(x=seq(0,50)) %>%
  mutate(pmf = map_dbl(x, ~dbinom(.x, size = 50, prob = 7/10)))

head(binom_df, 3)
```

## 1.3 (Q3)

```{r}
gaussian_df <- data.frame(x=seq(0,50,0.01)) %>%
  mutate(pdf = map_dbl(x, ~dnorm(.x, mean = 50*0.7, sd = sqrt(50*0.7*(1-0.7)))))

head(gaussian_df, 3)
```

## 1.3 (Q4)

```{r}
colors <- c("Gaussian pdf"="red", "Binomial pmf"="blue")
fill <- c("Gaussian pdf"="white", "Binomial pmf"="white")

ggplot() + labs(x="x", y="Probability") + theme_bw() +
  geom_line(data=gaussian_df, aes(x,y=pdf,color="Gaussian pdf"),size=2) +
  geom_col(data=binom_df, aes(x=x,y=pmf, color="Binomial pmf",fill="Binomial pmf")) +
  scale_color_manual(name="myLegend", values=colors) +
  scale_fill_manual(name="myLegend", values=fill) +
  xlim(c(20,50))
```

## 1.4 (Q1)

```{r}
f_musigma <- function(mu, sigma, x){
  y = (1/sigma/sqrt(2*pi))*exp(-0.5*((x-mu)/sigma)^2)
  return(y)
}

df_density <- data.frame(x=seq(-4,6,0.1)) %>%
  mutate('1' = dnorm(x=x,mean=1,sd=1),
         '2' = f_musigma(mu=1,sigma=sqrt(2),x=x),
         '3' = f_musigma(mu=1,sigma=sqrt(3),x=x))

df_density_longer = pivot_longer(df_density, col = c('1','2','3'),
                                 names_to = 'Variance', values_to = 'Density')

ggplot(df_density_longer, aes(x=x, y=Density, linetype=Variance, color=Variance)) +
  geom_line() + theme_bw()


```

## 1.4 (Q2)

```{r}
df <- data.frame(x=seq(-4,6,0.1)) %>%
  mutate('1' = pnorm(q=x, mean = 1, sd = 1),
         '2' = pnorm(q=x, mean = 1, sd = sqrt(2)),
         '3' = pnorm(q=x, mean = 1, sd = sqrt(3)))

df_longer = pivot_longer(df, col = c('1','2','3'),
                                 names_to = 'Variance', values_to = 'CDF')

ggplot(df_longer, aes(x=x, y=CDF, linetype=Variance, color=Variance)) +
  geom_line() + theme_bw()
```

## 1.4 (Q3)

```{r}
df_quantile <- data.frame(x=seq(0,1,0.005)) %>%
  mutate('1' = qnorm(p=x, mean = 1, sd = 1),
         '2' = qnorm(p=x, mean = 1, sd = sqrt(2)),
         '3' = qnorm(p=x, mean = 1, sd = sqrt(3)))

df_quantile_longer = pivot_longer(df_quantile, col = c('1','2','3'),
                                 names_to = 'Variance', values_to = 'Quantile')

ggplot(df_quantile_longer, aes(x=x, y=Quantile, linetype=Variance, color=Variance)) +
  geom_line() + theme_bw()
```

## 1.4 (Q4)

```{r}
set.seed(0)
standardGaussianSample = rnorm(100, mean = 0, sd = 1)
head(standardGaussianSample)
```

## 1.3 (Q5)

```{r}
mean1Var3GaussianSampleA = sqrt(3) * standardGaussianSample + 1
head(mean1Var3GaussianSampleA)
```

## 1.3 (Q6)

```{r}
set.seed(0)
mean1Var3GaussianSampleB = rnorm(100, mean = 1, sd = sqrt(3))
head(mean1Var3GaussianSampleB)

all.equal(mean1Var3GaussianSampleA, mean1Var3GaussianSampleB)
```

## 1.3 (Q7)

```{r}
colors <- c("Population density"="red", "Sample kernel density"="blue", "Population mean"="green", "Sample mean"="violet")
linetypes <- c("Population density"="solid", "Sample kernel density"="dashed", "Population mean"="solid", "Sample mean"="dashed")
 
ggplot() + labs(x="x",y="Density") + theme_bw() +
# 1. create plot of theoretical density
geom_line(data=select(df_density, x, "3"),
aes(x=x,y=`3`,color="Population density",
linetype="Population density")) +
# 2. add in kernel density plot from real sample
geom_density(data=data.frame(x=mean1Var3GaussianSampleA),
aes(x=x,color="Sample kernel density",
linetype="Sample kernel density")) +
# 3. vertical lines
geom_vline(aes(xintercept=1,color="Population mean",
linetype="Population mean")) +
geom_vline(aes(xintercept=mean(mean1Var3GaussianSampleA),
color="Sample mean",linetype="Sample mean")) +
# 4.
scale_color_manual(name = "Legend", values=colors) +
scale_linetype_manual(name="Legend", values=linetypes)
```

# 2. Location estimators with Gaussian data


## 2 (Q2)

```{r}
set.seed(0)
num_trials_per_sample_size <- 1000
min_sample_size <- 30
max_sample_size <- 500
sample_size_inc <- 5
mu_0 <- 1
sigma_0 <- 3
# create data frame of all pairs of sample_size and trial
simulation_df<-crossing(trial=seq(num_trials_per_sample_size),
sample_size=seq(min_sample_size,
max_sample_size,sample_size_inc)) %>%
# simulate sequences of Gaussian random variables
mutate(simulation=pmap(.l=list(trial,sample_size),
.f=~rnorm(.y,mean=mu_0,sd=sigma_0))) %>%
# compute the sample medians and sample means
mutate(sample_md=map_dbl(.x=simulation,.f=median), sample_mn=map_dbl(.x=simulation,.f=mean)) %>%
group_by(sample_size) %>%
summarise(msq_error_md=mean((sample_md-mu_0)^2), msq_error_mn=mean((sample_mn-mu_0)^2))

simulation_df %>%
  pivot_longer(cols = c(msq_error_md, msq_error_mn), names_to = "Estimator", values_to = "msq_error") %>%
  mutate(Estimator=case_when(Estimator=="msq_error_md"~"Median",
Estimator=="msq_error_mn"~"Mean")) %>%

ggplot(aes(x=sample_size, y=msq_error, color=Estimator, linetype=Estimator)) +
geom_smooth()+theme_bw()+xlab("Sample size")+ylab("Mean square error")
```

# 3. (**) The law of large numbers and Hoeffding’s inequality

## 3 (Q1)





