---
title: "Assignment3RMarkdown"
author: "Prithvi Abhishetty"
date: "2022-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

# 1. Exploratory data analysis


## 1.1 (Q1)   

```{r}
head(Hawks)
HawksTail = Hawks$Tail
head(HawksTail)
mean(HawksTail,na.rm = TRUE)
median(HawksTail,na.rm = TRUE)
```

## 1.2 (Q1)

```{r}
summarise(Hawks, Wing_mean=mean(Wing, na.rm=TRUE), Wing_t_mean=mean(Wing, trim=0.5, na.rm=TRUE),
Wing_med=median(Wing, na.rm=TRUE), Weight_mean=mean(Weight, na.rm=TRUE),
Weight_t_mean=mean(Weight, trim=0.5, na.rm=TRUE), Weight_med=median(Weight, na.rm=TRUE))
```


## 1.2 (Q2)

```{r}
group_by(Hawks, Species) %>%
summarise(Wing_mean=mean(Wing, na.rm=TRUE), Wing_t_mean=mean(Wing, trim=0.5, na.rm=TRUE),
Wing_med=median(Wing, na.rm=TRUE), Weight_mean=mean(Weight, na.rm=TRUE),
Weight_t_mean=mean(Weight, trim=0.5, na.rm=TRUE), Weight_med=median(Weight, na.rm=TRUE))
```
## 1.3 (Q1)

mean = a*A+b
```{r}
mean(HawksTail)*2+3
mean(HawksTail*2+3)
```
## 1.3 (Q2)

variance = aap
standard deviation = aq
```{r}
var(2*HawksTail+3, na.rm = TRUE)
2*2*var(HawksTail)

sd(2*HawksTail+3, na.rm = TRUE)
2*sd(HawksTail)
```
## 1.4 (Q1)

```{r}
hal<-Hawks$Hallux # Extract the vector of hallux lengths
hal<-hal[!is.na(hal)] # Remove any nans
outlier_val<-100
num_outliers<-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))

mean(hal)
mean(corrupted_hal)

num_outliers_vect <- seq(0,1000)
means_vect <- c()
for(num_outliers in num_outliers_vect){
corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
means_vect <- c(means_vect, mean(corrupted_hal))
}
head(means_vect)
```

```{r}
num_outliers_vect <- seq(0,1000)
medians_vect <- c()
for(num_outliers in num_outliers_vect){
corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
medians_vect <- c(medians_vect, median(corrupted_hal))
}
head(medians_vect)
```

## 1.4 (Q2)

```{r}
num_outliers_vect <- seq(0,1000)
t_means_vect <- c()
for(num_outliers in num_outliers_vect){
corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
t_means_vect <- c(t_means_vect, mean(corrupted_hal,trim = 0.1))
}
head(t_means_vect)
```

## 1.4 (Q3)

```{r}
df_means_medians <- data.frame(num_outliers=num_outliers_vect, mean=means_vect,
t_mean=t_means_vect, median=medians_vect)
head(df_means_medians)

df_means_medians %>%
pivot_longer(!num_outliers, names_to = "Estimator", values_to = "Value") %>%
ggplot(aes(x=num_outliers,color=Estimator, linetype=Estimator,y=Value)) +
geom_line()+xlab("Number of outliers")
```
the sample median is the most robust when the outliers are small.

## 1.5 (Q1)

```{r}
ggplot(data=Hawks, aes(x=Species, y=Weight))+geom_boxplot()+xlab('Species')+ylab("Weight")
```

## 1.5 (Q2)

```{r}
Hawks %>% 
  group_by(Species) %>% 
  summarise(quantile025=quantile(Weight, prob=0.25, na.rm = TRUE), quantile050=quantile(Weight, probs=0.5, na.rm=TRUE), quantile075=quantile(Weight, probs=0.75, na.rm=TRUE))
```

## 1.5 (Q3)

```{r}
num_outliers <- function(x){
  q25 <- x %>% quantile(0.25, na.rm = TRUE)
  q75 <- x %>% quantile(0.75, na.rm = TRUE)
  iq_range <- q75 - q25
  outliers <- sum((x > q75+1.5*iq_range)|(x < q25-1.5*iq_range), na.rm = TRUE)
  return (outliers)
}
num_outliers(c(0, 40, 60, 185 ))
```

## 1.5 (Q4)

```{r}
Hawks %>% group_by(Species) %>% summarise(num_outliers_weight=num_outliers(Weight))
```

## 1.6 (Q1)

```{r}
cov(Hawks$Weight,Hawks$Wing,use = 'complete.obs')
cor(Hawks$Weight,Hawks$Wing,use = 'complete.obs')
```


## 1.6 (Q2)

covariance = acS
correlation = sign(ac)R (not sure why)
```{r}
2.4*-1*cov(Hawks$Weight,Hawks$Wing,use = 'complete.obs')
cov(Hawks$Weight*2.4+7.1, Hawks$Wing*(-1)+3,use = 'complete.obs')

sign(2.4*-1)*cor(Hawks$Weight,Hawks$Wing,use = 'complete.obs')
cor(Hawks$Weight*2.4+7.1, Hawks$Wing*(-1)+3,use = 'complete.obs')

```


# 2. Random experiments, events and sample spaces, and the set theory

## 2.1 (Q1)
Random Experiment: has a well-defined set of possible outcomes. Could be repeated arbitrarily many times.
Event: A set of possible outcomes of an experiment
Sample Space: The set of all possible outcomes of interest for a random experiment

## 2.2 (Q2)

Event example: {(1,2),(2,3)}
Sample space: {(a,b) | a,b ∈ {1,2,3,4,5,6}}
Total number of different events: 36 possible outcomes so 2^(36) different events.
The empty set is an event.

## 2.2 (Q1)

1. {1,2,3,4,6}, {1,2,3,4,5,6}
2. {2}, {∅}
3. {1,3}, {1,2,3}
4. No, Yes
5. Yes
6. {1,2},{3,4,5,6}; {1},{2},{3,4,5,6}

## 2.2 (Q2)

1. A
2. {∅}
3. if A ⊆ B, then A<B. (B^c = 1-B) < (1-A = A^c) therefore B^c ⊆ A^c.
4. long
5. long
6. long

## 2.2 (Q3)

2^K

## 2.2 (Q4)

1. The empty set
2. long

## 2.2 (Q5)

1. 1-1A
2. Ω = A ∪ A^c
3. long

## 2.2 (Q6)

long

# 3. Visualisation

## 3 (Q1)

```{r}
ggplot(Hawks, aes(x=Tail, color=Species))+geom_density()+theme_bw()+xlab("Tail (mm)")+ylab("Density")
```

## 3 (Q2)

```{r}
ggplot(Hawks, aes(x=Tail, y=Species, fill=Species))+geom_violin()+theme_bw()+xlab("Tail (mm)")+ylab("Density")
```

## 3 (Q3)

```{r}
ggplot(Hawks, aes(y=Weight,x=Tail))+xlab("Tail (mm)")+ylab("Weight (mm)")+geom_point(aes(color=Species, shape=Species))+theme_bw()
```
Four aesthetics: Tail length is mapped to horizontal position, Weight is mapped to vertical position, 
Species are mapped to colour, Species are mapped to shape.
The glyphs are the small elements (represented by the shapes) corresponding to individual cases.
Visual cues include horizontal and vertical position, shape, colour.

## 3 (Q4)

```{r}
ggplot(Hawks, aes(y=Weight,x=Tail))+xlab("Tail (mm)")+ylab("Weight (mm)")+geom_point(aes(color=Species, shape=Species))+theme_bw()+facet_wrap(~Species, scales='free')+geom_smooth(method = 'lm')
```
Visual cues include the horizontal and vertical positions, colour, shape, direction.
Based on this sample, longer tail lengths appear to be predictive of larger weights within each species.

## 3 (Q5)

```{r}
max_weight<-max(Hawks$Weight, na.rm=TRUE)
max_weight_row = head(Hawks %>% filter(Weight>=max_weight), 1)
max_weight_tail = max_weight_row$Tail

ggplot(Hawks, aes(y=Weight,x=Tail))+xlab("Tail (mm)")+ylab("Weight (mm)")+geom_point(aes(color=Species, shape=Species))+theme_bw()+geom_curve(x=max_weight_tail,xend=max_weight_tail,y=max_weight-200,yend=max_weight,
arrow=arrow(length=unit(0.1,"cm")), color="black")+
geom_text(x=max_weight_tail,y=max_weight-200,label='heaviest hawk',size=3,color="black")
```
